---
title: "Reproducible Research Project 1"
author: "Sujay Nair"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1.
Reading in the data/getting a feel for the data

```{r}
activitydata <- read.csv("activity.csv")
print(summary(activitydata))
print(head(activitydata))
```

##2.
Creating a histogram for the total number of steps taken each day and finding the mean and median of the total number of steps

```{r}
totalsteps <- aggregate(activitydata$steps, list(activitydata$date), sum)
print(head(totalsteps))
hist(totalsteps$x, xlab = "Number of Steps", main = "Steps per Day")
print(mean(totalsteps$x, na.rm = TRUE))
print(median(totalsteps$x, na.rm = TRUE))
```

##3.
Finding the average daily activity pattern using a time series plot

```{r}
averagesteps <- aggregate(activitydata$steps, list(factor(activitydata$interval))
                          , mean, na.rm = TRUE)
print(head(averagesteps))
plot(averagesteps$Group.1, averagesteps$x, type = "l")
#Here we print the row with the maximum average value to see the interval with the max number of steps
print(averagesteps[which.max(averagesteps$x),])
```

##4.
Working with missing values

```{r}
#Total number of missing values
NAs <- (is.na(activitydata))
NAdataset <- activitydata[NAs,]
#I filled in the NAs with the mean at each interval
avgsteps_perint <- aggregate(activitydata$steps, list(activitydata$interval), mean, na.rm = TRUE)
print(head(avgsteps_perint))
colnames(avgsteps_perint)[colnames(avgsteps_perint)=="Group.1"] <- "interval"
colnames(avgsteps_perint)[colnames(avgsteps_perint)=="x"] <- "average_value"
mergeddataset <- merge(activitydata, avgsteps_perint, by = "interval")
moreNAs <- is.na(mergeddataset$steps)
mergeddataset[moreNAs,] = avgsteps_perint[moreNAs,]
print(head(mergeddataset))
newtotalsteps <- aggregate(mergeddataset$steps, list(mergeddataset$date), sum)
print(head(newtotalsteps))
hist(newtotalsteps$x, col = "purple")
```



##5
Checking for differences in activity patterns in weekdays verse weekends

```{r}
activitydata$date <- as.POSIXct(strptime(activitydata$date, "%Y-%m-%d"))
activitydata$day <- 0
print(length(activitydata$date))
for(i in 1:length(activitydata$date)){
    if(weekdays(activitydata$date[i]) %in% c("Sunday", "Saturday")){
      activitydata$day[i] = "weekend"
    } else{
      activitydata$day[i] = "weekday"
    }
  }
print(unique(activitydata$day))

weekendDAYS <- subset(activitydata, activitydata$day == "weekend")
weekdayDAYS <- subset(activitydata, activitydata$day == "weekday")
WKNDaveragesteps <- aggregate(weekendDAYS$steps, list((weekendDAYS$interval))
                          , mean, na.rm = TRUE)
WKDYaveragesteps <- aggregate(weekdayDAYS$steps, list((weekdayDAYS$interval))
                          , mean, na.rm = TRUE)
par(mfrow=c(2,1))
plot(WKDYaveragesteps$Group.1, WKDYaveragesteps$x, type = "l", col = "blue", xlab = "Interval"
     ,ylab = "Steps", main = "AVG Steps(Weekdays)", ylim = c(0,250))
plot(WKNDaveragesteps$Group.1, WKNDaveragesteps$x, type = "l", col = "red", xlab = "Interval"
     ,ylab = "Steps", main = "AVG Steps(Weekends)", ylim = c(0,250))


